---
title: "2.viewing"
output: 
  html_document:
    toc: yes
---

# Viewing and understanding the training data

## Loading the data

We load from the compiled, preformatted data saved earlier :

```{r Loading environement}
load("d.train.RData")
load("im.train.RData")
load("im.test.RData")
ls()
```

## Basic quantitative summaries


```{r}
dim(d.train)
dim(im.train)

dim(im.test)

summary(d.train, digits = 1)[c(1,3,4,6),]

```

## Missing analysis


As we can see below, __only the nose_tip__ is garanteed to be present all the time.
A few eye centers (13) and mouth centers (33) are missing, but most are present.
A lot of the rest (~ 4800, i.e. ~ 70% ) are missing !

```{r}
as.matrix(summary(d.train, digits = 1)[7,])
```


Let's check if x and y missing always match, as one would expect. More specifically, we are looking for a situation where either the x or the y value is NA, but not both, for the same keypoint.

```{r}
any(xor(is.na(d.train[, seq(1,30,2)]),is.na(d.train[, seq(2,30,2)])))
```

Good. There is none.




## Plotting utility functions

Plotting the keypoints, as an overlay to an existing earlier image.

__Keys represent a dataframe where keys are the rows. Coordinates need to be adjusted, because matrix origin is top left, while display origin is bottom left.__

```{r}
m.plotKeypoints <- function(keys, col="red") {
  x <- 96 - as.numeric(keys[seq(1,30,2)])
  y <- 96 - as.numeric(keys[seq(2,30,2)])
  points(x,y,col=col)
  lines(mean(x),mean(y),col="yellow",pch=3)
  # x value for various parts
  le <- c(5,1,7) # left eye 
  re <- c(11,3,9) #  right eye 
  lb <- c(13,15) # left eyBrow
  rb <- c(17,19) # right eyeBrow
  mm <- c(23,27,25,29,23) # mouth
  
  lines(x = 96-keys[le],y=96-keys[le+1],col=col)
  lines(x = 96-keys[re],y=96-keys[re+1],col=col)
  lines(x = 96-keys[lb],y=96-keys[lb+1],col=col)
  lines(x = 96-keys[rb],y=96-keys[rb+1],col=col)
  lines(x = 96-keys[mm],y=96-keys[mm+1],col=col)
  
}
```

Plotting an image, optionnally with a title and keypoints.

__The image is a row 1 x 9296.__

```{r}
m.plotImage <- function(image=rep.int(128,9216), kp=NA , title=NA, col="red") {
  n <- sqrt(length(image))
  image(
    1:n,
    1:n,
    matrix(rev(image), nrow=n, ncol=n),
    col=gray((0:255)/255),asp=1.)
  if(!any(is.na(kp))) {
    m.plotKeypoints(kp,col)
  }
  if(!is.na(title)) {
    title(main=title)
  }
}
```

This is how it works : 

```{r}
m.plotImage(im.train[1,],kp = d.train[1,])
m.plotImage(im.train[1,],kp = d.train[1,],col="orange")
```

## Symetries


The function m.mirror will produce a symetrical key w.r.t. the middle vertical axis, by replacing all left element with their right element, while mirorring all coordinates.

```{r}
m.mirrorKey <- function(key) {
  r <- key
  r[seq(1,30,2)] <- 96-key[seq(1,30,2)]
  left <-  c(1,5,7,13,15,23 ,    2,6,8,14,16,24)
  right <- c(3,9,11,17,19,25,    4,10,12,18,20,26 ) 
  a <- r[right]
  r[right] <- r[left]
  r[left] <- a
  return (r)
}
```

Let's see on an example

```{r}
m.plotImage(im.train[22,], kp=d.train[22,], col="red")
m.plotKeypoints(m.mirrorKey(d.train[22,]),col="blue")
title(main="Original(red) vs. mirror(blue)")
```

## Compute average figure

Let's create the "average" figure from the complete training dataset.

```{r}
d.key.mean <- 
  colMeans(na.omit(d.train))
  length(d.key.mean)
```

Let's compare this with the means we get using all non NA cell, even on incomplete lines.

```{r}
d.key.mean.all <- colMeans(d.train, na.rm = TRUE)
length(d.key.mean.all)

plot(x=d.key.mean, y=d.key.mean.all )
title(main="Mean all vs. rigourous mean")
lines(x=c(0,95),y=c(0,95),col="yellow")

sort(d.key.mean - d.key.mean.all)[-5:-26]
```

We see that there are only significant differences, when computing the means in both ways, for **"nose_tip_y"** and for **"mouth_center_bottom_lip_y"**

Graphically :


```{r}
m.plotImage()
m.plotKeypoints(d.key.mean,col="red")         # Moyenne sur elements strictement complets
m.plotKeypoints(d.key.mean.all,col="blue")    # Moyenne extensive
lines(x=c(95/2,95/2),y=c(1,96),col="yellow")  # Axe de symetrie latérale
title(main="Compare d.key.mean vs. d.key.mean.all")
```

Let's check if we have symetry issues here ?

```{r}
m.plotImage()
m.plotKeypoints(d.key.mean, col="red")
m.plotKeypoints(m.mirrorKey(d.key.mean), col="green")
title(main="d.key.mean and mirror")

m.plotImage()
m.plotKeypoints(d.key.mean.all, col="red")
m.plotKeypoints(m.mirrorKey(d.key.mean.all), col="green")
title(main="d.key.mean.all and mirror")

```


There does not seem to be any noticeable difference by symetry.

**We therefore decide to __adjust__ both means for symetry**.

```{r}

m.key.mean <- (d.key.mean + m.mirrorKey(d.key.mean))/2
m.key.mean.all <- (d.key.mean.all + m.mirrorKey(d.key.mean.all))/2


m.plotImage()
m.plotKeypoints(d.key.mean, col="red")
m.plotKeypoints(m.key.mean, col="yellow")
title(main="d.key.mean vs. m.k.mean(yellow)")

m.plotImage()
m.plotKeypoints(d.key.mean.all, col="red")
m.plotKeypoints(m.key.mean.all, col="yellow")
title(main="d.key.mean.all vs. m.k.mean.all(yellow)")



```


## Considering adjusting missing elements ?

The massively missing elements, were we have only the nose, should probably not be adjusted. The rest is neglectable, and should probably be discarded.

For the time being, we would just focus on the complete examples :

```{r}
d.train.complete <- which(!is.na(rowSums(d.train)))
```

There are `r length(d.train.complete)` complete training records, out of the `r dim(d.train)[1]` total training records.

## Cleaning and saving

We save selectively what we changed or created.

```{r Saving environment}
save(d.train,d.key.mean,d.key.mean.all, m.key.mean,m.key.mean.all, d.train.complete,file="d.train.RData")
save(m.plotImage, m.plotKeypoints,m.mirrorKey,file="functions.viewing.RData")
```


*This file was compiled on  `r date()`*
